%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  08 April 2002	Version 4				    					%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%								    								%
%%  Doctoral Dissertation By Will Zhang at							%
%%	The University of Texas at Austin			     				%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[12pt]{report}

\usepackage{utdiss2}	% The dissertation style as required by UT.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Optional packages used for this dissertation. 					%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{amsmath,amsthm,amsfonts,amscd,commath,mathtools} % Mathematic fonts
%%. and styles used in this dissertation
\DeclareMathAlphabet\mathbfcal{OMS}{cmsy}{b}{n}	% Bolded caligraphy font

\usepackage{verbatim}      	% Allows quoting source with commands.
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\graphicspath{ {images/} }

\usepackage{array,multirow,graphicx}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}


\usepackage{etoolbox}
\apptocmd{\sloppy}{\hbadness 10000\relax}{}{}
\apptocmd{\thebibliography}{\raggedright}{}{}

\usepackage{enumitem}

\usepackage[table]{xcolor}
\definecolor{Gray}{gray}{0.9}

% Used to create the nomenclatures					%
\makeatletter

\newlength\tdima
\newcommand\tabfill[1]{%
      \setlength\tdima{\linewidth}%
      \addtolength\tdima{\@totalleftmargin}%
      \addtolength\tdima{-\dimen\@curtab}%
      \parbox[t]{\tdima}{#1\ifhmode\strut\fi}}

      
\newcommand\mytabs{\hspace*{0.5in}\=\hspace*{0.25in}\=\hspace*{0.25in}\=\hspace*{0.25in}\=\hspace*{0.25in}\=\hspace*{0.25in}\=\hspace*{0.25in}\=\hspace*{0.25in}\=\hspace*{0.25in}\=\hspace*{0.25in}\=\hspace*{0.25in}\=\hspace*{0.25in}}
\newenvironment{mynom}[1][\mytabs]
  {\begin{tabbing}#1\kill\ignorespaces}
  {\end{tabbing}}
% Used to create the nomenclatures					%
  
  
\usepackage{rotating}

\usepackage{eucal} 	 	% Euler fonts
\usepackage{makeidx}       	% Package to make an index.
\usepackage{url}		% Allows good typesetting of web URLs.

\usepackage{extdash}

\usepackage[square,numbers,sectionbib]{natbib}
\usepackage{chapterbib}

\usepackage{bibentry}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  About the Authors										%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\author{Will Zhang}  	% Author

\previousdegrees{B.S.}

\address{13926 Winding Cypress Brook Dr. \\ Cypress, Texas 77429}  % Address

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  About the dissertation									%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Modeling and simulating time-dependent changes in soft-tissue fabricated bioprosthetic heart valve in response to cyclic loading}

\supervisor
	{Michael S. Sacks}
    
\committeemembers
	[Aaron Baker]
	[Christopher G. Rylander]
	[Krishnaswamy Ravi-Chandar]
	{Naren Vyavahare}
    

\oneandonehalfspacequote

\topmargin 0.1in


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  Macros that are needed									%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\latexe}{{\LaTeX\kern.125em2%
                      \lower.5ex\hbox{$\varepsilon$}}}

\newcommand{\amslatex}{\AmS-\LaTeX{}}

\chardef\bslash=`\\	% \bslash makes a backslash (in tt fonts)
			%	p. 424, TeXbook

\newcommand{\cn}[1]{\texttt{\bslash #1}}

\makeatletter		% Starts section where @ is considered a letter
			% and thus may be used in commands.
\def\square{\RIfM@\bgroup\else$\bgroup\aftergroup$\fi
  \vcenter{\hrule\hbox{\vrule\@height.6em\kern.6em\vrule}%
                                              \hrule}\egroup}
\makeatother		% Ends sections where @ is considered a letter.
			% Now @ cannot be used in commands.

\makeindex    % Make the index
 
 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Beginning of the document										     %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
 
\begin{document}

\copyrightpage          % Produces the copyright page.

% NOTE: In a doctoral dissertation, the Committee Certification page
%		(with signatures) is BEFORE the Title page.
%	In a masters thesis or report, the Signature page
%		(with signatures) is AFTER the Title page.
%
%	If you are writing a masters thesis or report, you MUST REVERSE
%	the order of the \commcertpage and \titlepage commands below.
%
\commcertpage           % Produces the Committee Certification
			%   of Approved Version page (doctoral)
			%   or Signature page (masters).
			%		20 Mar 2002	cwm

\titlepage              % Produces the title page.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Dedication and/or epigraph are optional, but must occur here.      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% \begin{dedication}
% \index{Dedication@\emph{Dedication}}%
% Dedicated to my wife Shirley.
% \end{dedication}

\begin{acknowledgments}		% Optional
\index{Acknowledgments@\emph{Acknowledgments}}%
    Firstly, I would like to thank my advisor, Dr. Michael Sacks for the introducing to me to the world of biomechanics, the opportunity to work with his research group, and his patience and guidance over the years to prepare me for an academic career. Also, thank you to my committee members andthe Department of Biomedical Engineering at the University of Texas at Austin for their support throughout this process. I would also like to thank. It has been an honor to represent this department in my time here.
    

    I would like to thank my wonderful lab mates for giving me strength and support throughout this long process, lifting me up when I needed it the most. I also like to thank the post-doctoral fellows in our lab in particular, specifically Rong Fan, Chung-Hao Lee, Ankush Aggarwal and Joao Soares for their mentorship and guidance both in research and academically.
    

    Lastly, and most importantly I would like to thank my family, my parents, for raising me to the way I am, and all of the support they have given me throughout my life. I love you all.
\end{acknowledgments}

\utabstract
\index{Abstract}%
\indent
Soft-tissue-derived exogenously cross-linked (EXL) biomaterials continue to the best choice for the fabrication of bioprosthetic heart valves (BHV). Despite the ongoing research, our understanding of these materials and of the mechanisms leading to their failure remain at an empirical level. The need for advancements in modeling their behavior is further underscored by the development of percutaneously-delivered BHV devices. While these devices offer reduced surgical risk, they also present additional challenges for the design of the leaflets due to limitations in thickness and folding during delivery, resulting in a 2-year mortality rate of 33.9\% in general[2]. Thus, we seek to develop a framework for modeling and simulating soft-tissue-derived EXL biomaterials, accounting for the effects of exogenous cross-linking in permanent set and mechanical fatigue. Such approaches can significantly improve the accuracy and reliability of long-term predictions of durability and mechanical function. Firstly, we will establish the form of a nonlinear hyperelastic meso-scale structural constitutive model (MSSCM) for fibrous soft tissues, that can accurately capturing the mechanical response of common soft tissues used as a basis for EXL biomaterial. Secondly, we will study the effect of exogenous crosslinks on these tissues using glutaraldehyde (GLUT) EXL bovine pericardium. GLUT-EXLs form polymeric chains through the cross-linking process which more tightly bonds the fibers to the matrix, increasing the non-fibrous matrix stiffness and fiber-fiber interactions.  However, GLUT EXLs undergo Schiff-base reactions that lead to scission-healing behaviors that changes the geometry of BHVs. We model this effect based on first order kinetics of the scission healing reaction and validate it using static strain, cyclic strain and stress control experiments. Next, we will develop a full 3D finite element implementation of the MSSCM with the modifications for EXL for real device applications. We then parametrically examine the changes in geometry and stress distribution of BHVs overtime, exploring initial geometries and material properties which may minimize the risks of the former effects. With this, we aim to develop a better understanding of the underlying process that occurs during long-term cyclic loading through our constitutive modeling approach and device level applications, and translate the insights gained to improve BHV design and durability. 
 
\tableofcontents{}

\listoftables      % List of Tables and List of Figures will be placed
\listoffigures     % here, if applicable.

\include{Chapters/chapter1-introduction}

\include{Chapters/chapter2-constitutive}

%\include{Chapters/chapter3-biaxialtesting}

% \include{Chapters/chapter3-softtissue}

\include{Chapters/chapter3-crosslinking}

\include{Chapters/chapter4-permanentset}

\include{Chapters/chapter5-effectivemodel}

\include{Chapters/chapter6-simulation}


\appendices

% \include{Chapters/appendixA}

\include{Chapters/appendixB}


%\bibliographystyle{plain}  % Here the bibliography 		     %
%\bibliography{phd}        % is inserted.			     %
%\index{Bibliography@\emph{Bibliography}}%			     %


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Generate the index.						     %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%								     %
% NOTE: For master's theses and reports, NOTHING is permitted to     %
%	come between the bibliography and the vita. This section     %
%	to generate the index (if used) MUST be moved to before      %
%	the bibliography section.				     %
%								     %
%\printindex%    % Include the index here. Comment out this line      %
%		% with a percent sign if you do not want an index.   %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Vita page.							     %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{vita}
Will Zhang, 
born in Linhe, Inner Mongolia, China, on 10 August 1989, 
the son of Dr. Wenxiu Zhang and Guiping Tian.  He received the Bachelor
of Science in Honours Biophysics from University of British Columbia.
Continuing his academic career, he applied to the University of Texas at Austin for enrollment in their graduate program in Biomedical Engineering. He was accepted and started graduate studies in May, 2018.

\end{vita}


\end{document}


