\chapter{Effective model representation of structural and multi-scale models using effective constitutive models for facilitating numerical simulations}

\section*{Preface}
\addcontentsline{toc}{section}{Preface}%

One of the most crucial aspects of biomechanical simulations of organs and systems that seek to predict the outcomes of disease, injury, and surgical interventions is the underlying constitutive model. Current soft tissue constitutive modeling approaches have become increasingly complex, often utilizing meso- and multi-scale methods for greater predictive capability and linking to the underlying mechanisms. However, such modeling approaches are associated with substantial computational costs. One solution is to use effective constitutive models, which only reproduces the essential responses but not the underlying mechanisms. Effective constitutive models can be implemented in place of meso- and multi-scale models in numerical simulations, but derive their responses by homogenizing the responses of the underlying meso- or multi-scale models. A robust effective constitutive model can thus drastically increase the speed of simulations for a wide range of meso- and multi-scale models. However, there is no general consensus on how to develop a single effective constitutive model for a wide range of soft tissue responses. In the present study, we developed an effective constitutive model, which can fully reproduce the response of a wide range of planar soft tissue responses, along with methods for fast-convergent parameter estimation. We evaluated this approach and demonstrated that it is able to handle materials of widely varying degrees of anisotropy, such as exogenously cross-linked bovine pericardium and aortic valve leaflet. This effective constitutive model approach has shown significant potential for improving the computational efficiency and numerical robustness of multi-scale and meso-scale modeling approaches, facilitating the application of inverse modeling and simulations of growth and remodeling of soft tissues and organs.

% \textbf{The work contained in this chapter was published as}: Zhang, W., Zakerzadeh, Z., Zhang, W. \& Sacks, M. S.
% A Material Modeling Approach for the Effective Response of Planar Soft Tissues for Efficient Computational Simulations. 
% Journal of the mechanical behavior of biomedical materials. under review.



%---    INTRODUCTION
\input{Chapters/chapter5-content/chapter5.1.tex}

%---    METHODS
\input{Chapters/chapter5-content/chapter5.2.tex}

%---    Results
\input{Chapters/chapter5-content/chapter5.3.tex}

%---    Discussion
\input{Chapters/chapter5-content/chapter5.4.tex}

%---    Conclusion
\input{Chapters/chapter5-content/chapter5.5.tex}



\newpage

%---    Bioliography
\bibliographystyle{plainnat}
\bibliography{phd}