\chapter{Effect of exogenous cross-linking on the mechanical response of soft tissues}


\section*{Preface}
\addcontentsline{toc}{section}{Preface}%

    Exogenous cross-linking of soft collagenous tissues is a common method for biomaterial development and medical therapies. It is an important part of the fabrication process of BHVs and significantly affects their mechanical properties. To enable improved applications through computational methods, physically realistic constitutive models are required. Yet, despite decades of research, development and clinical use, no such model exists. In this study, we develop the first rigorous full structural model (i.e. explicitly incorporating various features of the collagen fiber architecture) for exogenously cross-linked soft tissues. This was made possible, in part, with the use of native to cross-linked matched experimental datasets and an extension to the collagenous structural constitutive model so that the uncross-linked collagen fiber responses could be mapped to the cross-linked configuration. This allowed us to separate the effects of cross-linking from kinematic changes induced in the cross-linking process, which in turn allowed the non-fibrous tissue matrix component and the interaction effects to be identified. It was determined that the matrix could be modeled as an isotropic material using a modified Yeoh model. The most novel findings of this study were that: (i) the effective collagen fiber modulus was unaffected by cross-linking and (ii) fiber-ensemble interactions played a large role in stress development, often dominating the total tissue response (depending on the stress component and loading path considered). An important utility of the present model is its ability to separate the effects of exogenous cross-linking on the fibers from changes due to the matrix. Applications of this approach include the utilization in the design of novel chemical treatments to produce specific mechanical responses and the study of fatigue damage in bioprosthetic heart valve biomaterials.

\textbf{The work contained in this chapter was published as}:  Sacks, M. S.; Zhang, W. \& Wognum, S.
A novel fibre-ensemble level constitutive model for exogenous cross-linked collagenous tissues 
Interface Focus, The Royal Society, 2016, 6, 20150090 


%---    INTRODUCTION
\input{Chapters/chapter3-content/chapter3.1.tex}

%---    Experimental methods and data post-processing
\input{Chapters/chapter3-content/chapter3.2.tex}

%---    Delineation and modelling of the tissue-level mechanical effects of exogenous cross-links
\input{Chapters/chapter3-content/chapter3.3.tex}

%---    Initial model formulation
\input{Chapters/chapter3-content/chapter3.4.tex}

%---    Model simplifications and parameter estimation
\input{Chapters/chapter3-content/chapter3.5.tex}

%---    primary results
\input{Chapters/chapter3-content/chapter3.6.tex}

%---    discussion
\input{Chapters/chapter3-content/chapter3.7.tex}

\newpage
\input{Chapters/chapter3-content/nomenclature.tex}

\newpage
%---    Bioliography
\bibliographystyle{plainnat}
\bibliography{phd}