\chapter{Numerical simulation of intact bioprosthetic heart valves}

\section*{Preface}
\addcontentsline{toc}{section}{Preface}%

    Bioprosthetic heart valves (BHVs), fabricated from exogenously crosslinked collagenous tissues, remain the most popular heart valve replacement design. However, the life span of BHVs remains limited to 10–15 years, in part because the mechanisms that underlie BHV failure remain poorly understood. The current process for evaluating BHV designs is an expensive and time consuming three-stage process: 1) accelerated wear testing(AWT), 2) large animal studies, and 3) clinical trials. AWT is currently the only way to evaluate BHV durability in a feasible amount of time (months instead of years). However, the loading conditions and environment during AWT are not physiological and the only durability information currently used is the presence of visible damage. Only the clinical evaluation stage can provide true indications of the \textit{in vivo} performance of BHV designs. However, this is the final, most difficult, most expensive and most time consuming stage. Clearly, current methods of evaluating BHV designs are not feasible for advancing the BHV technology in a timely manner. In this chapter, we implement our constitutive model for the time evolving properties of BHVs in response to permanent set developed in a previous chapter in a numerical simulation framework, and attempt to predict the geometric and structural change that occurs with the onset of permanent set. 




%---    INTRODUCTION
\input{Chapters/chapter6-content/chapter6.1.tex}

%---    METHODS
\input{Chapters/chapter6-content/chapter6.2.tex}

%---    Results
\input{Chapters/chapter6-content/chapter6.3.tex}

%---    Discussion
\input{Chapters/chapter6-content/chapter6.4.tex}




%---    Bioliography
\bibliographystyle{plainnat}
\bibliography{phd}